### 2016-06-29 paper/abstract.tex

Tools for reliably converting back and forth between different data
formats are increasingly important in a world full of big data.
One class of such tools are \emph{lenses}, programs that simultaneously
represent one function to convert data from type $A$ to type $B$ and a
second function to convert data from type $B$ back into data of type
$A$.  While such programs often come with important invertability properties,
they can be quite difficult to write in practice 
as a programmer must think about data conversions in two directions as once.
To overcome this difficulty, we develop algorithms for synthesizing such lenses 
automatically from their types and a small number of examples.
Unlike past work in type-directed synthesis, synthesis of lenses is
complicated by their rich type structure, which is based on regular
expressions.  More specifically, rather than simply having to solve
the problem of finding a representative term to map back and forth
between types $A$ and $B$, it is first necessary to find the right
representative types $A'$ and $B'$ amongst many equivalent regular expressions.
We show how to manage this search by defining a new language of
lenses governed by semi-canonical regular expression types and prove it
sound with respect to a standard language of bijective lenses.  We then show
how to search the reduced space for well-typed lenses that are consistent
with the given examples.  We demonstrate the effectiveness of our algorithms
by experimenting on benchmark suite containing microbenchmarks, FlashFill
examples, and linux configuration file transformations.

>>>>

The growing dependence on big data makes reliably converting
between different data formats increasingly important.
\emph{Lenses} are a bi-directional technique for specifying such
conversions.  Specifically, a lens is a declarative program that
simultaneously specifies a function to convert data from type $A$ to
type $B$ and a second function to convert data from type $B$ to type
$A$.  Such programs usually provide invertability guarantees, avoiding
pernicious bugs that can corrupt data and downstream processing.
Unfortunately, lenses can be quite difficult to write because they
require programmers to think in two directions at once.  To overcome
this difficulty, we develop an algorithm to automatically synthesize
lenses from format descriptions, specified as regular expressions, and
a small number of examples.  The rich structure of regular expressions
makes lens synthesis more complicated than previous work on other
forms of type-directed synthesis.  Specifically, finding a term to
bi-directionally map between regular expressions $A$ and $B$ requires
first finding equivalent regular expressions $A'$ and $B'$ that are
more algorithmically tractable.
We show how to manage this search by defining a new language of
lenses governed by semi-canonical regular expression types and prove it
sound with respect to a standard language of bijective lenses.  
We then show
how to search the reduced space for well-typed lenses that are consistent
with the given examples.  We demonstrate the effectiveness of our algorithm
by experimenting on a benchmark suite containing XX examples
comprising microbenchmarks, FlashFill
examples, and Augeus transformations of Linux configuration files.
We show we are able to infer lenses for most examples in the
benchmark suite in an average of XXX seconds.
